Threshold Phenomenon in a Nonlinear Heat Equation

### 2016-04-22 pde_two/threshold.tex
yolo
almost done :D

This paper concerns the behavior of the following non-linear heat equation given by
\begin{gather}
u_{t}=u_{xx}+u^{3},\ 0<x<\pi,\ t>0\\
u(0,t)=u(\pi,t)=0,\ t>0\\
u(x,0)=u_{0}(x),\ 0<x<\pi.
\end{gather}

@@

\section{Steady State Analysis}
The steady state equation of $(1)$ is given by
\begin{gather}
u_{xx}+u^{3}=0\\
u(0)=u(\pi)=0.
\end{gather}
We want to find a solution $u_{s}(x)$ that satisfies the following properties
\begin{enumerate}
\item If $u_{0}(x)<u_{s}(x)$, then $u(x,t)\to u_{s}(x)$ as $t\to\infty$.
\[
\lim_{t\to \infty}u(x,t)=u_{s}(x).
\]
In this case, the solution to $(1)$ would independent of time. If this property is fulfilled, then
there is a \textit{global solution} $u(x,t)$ to the boundary value problem described in $(1)$.
\item If $u_{0}(x)>u_{s}(x)$, then $u(x,t)$ blows up in a finite time. This means that the solution
tends towards infinity in a finite time $T$.
\[
\lim_{t\to T}u(x,t)=\infty.
\]
This case if often indicative of a singularity in the solution $u(x,t)$. The solution $u$ of $(1)$
would not be global in this case.
\end{enumerate}
We can now attempt to solve this steady state problem.

@@

If we can find the average temperature of the steady state, then we know that our function $u(x)$
has a global solution (otherwise there would be points in the range of $u$ for which $u$ blows up).
We will want to show that
\[
{1\over\pi}\int_{0}^{\pi}u(x)\dx={\sqrt{2}\over2}.
\]
To begin, let
\[
y=\int_{0}^{x}u(t)\dt
\]
and $v(y)=u^{2}(x)$ where $x=x(y)=y^{-1}(x)$.

\end{document}

>>>>

This paper concerns the behavior of the following non-linear heat equation given by
\begin{gather}
u_{t}=u_{xx}+u^{3},\ 0<x<\pi,\ t>0\label{nlinh}\\
u(0,t)=u(\pi,t)=0,\ t>0\\
u(x,0)=u_{0}(x),\ 0<x<\pi.
\end{gather}

@@

\section{Steady State Analysis}
The steady state equation of $(1)$ is given by
\begin{gather}
u_{xx}+u^{3}=0\label{ssone}\\
u(0)=u(\pi)=0.
\end{gather}
We want to find a solution $u_{s}(x)$ that satisfies the following properties
\begin{enumerate}
\item If $u_{0}(x)<u_{s}(x)$, then $u(x,t)\to u_{s}(x)$ as $t\to\infty$.
\[
\lim_{t\to \infty}u(x,t)=u_{s}(x).
\]
In this case, the solution to $(1)$ would independent of time. If this property is fulfilled, then
there is a \textit{global solution} $u(x,t)$ to the boundary value problem described in $(1)$.
\item If $u_{0}(x)>u_{s}(x)$, then $u(x,t)$ blows up in a finite time. This means that the solution
tends towards infinity in a finite time $T$.
\[
\lim_{t\to T}u(x,t)=\infty.
\]
This case if often indicative of a singularity in the solution $u(x,t)$. The solution $u$ of $(1)$
would not be global in this case.
\end{enumerate}
We can now attempt to solve this steady state problem.

@@

If we can find the average temperature of the steady state, then we know that our function $u(x)$
has a global solution (otherwise there would be points in the range of $u$ for which $u$ blows up).
We will want to show that
\[
{1\over\pi}\int_{0}^{\pi}u(x)\dx={\sqrt{2}\over2}.
\]
To begin, let
\[
y=\int_{0}^{x}u(t)\dt \implies {\dy\over\dx} = u(x)
\]
and $v(y)=u^{2}(x)$ where $x=x(y)=y^{-1}(x)$. Finally, we will let
\[
R=\int_{0}^{\pi}u(x)\dx.
\]
We now can set up an ordinary differential equation in terms of $v$.
\begin{align*}
v'(y)&=2u(x)u'(x){\dx\over\dy}\\
&= 2{\dy\over\dx}u'(x){\dx\over\dy}\\
&= 2u'(x)\\
v''(y)&= 2u''(x){\dx\over\dy}\\
&= {2u'(x)\over u(x)}.
\end{align*}
This implies that $v$ satisfies the following linear ordinary differential equation:
\begin{gather}\label{vode}
v''(y)+2v(y)=0
\end{gather}
on the interval $0<y<R$ where $v(0)=v(R)=0$. The differential equation in \eqref{vode} has the
characteristic equation
\[
r^{2}+2=0,
\]
implying that the basis of our solution is the in the form $\{\sin{\sqrt{2}y},\cos{\sqrt{2}y}\}$.
Applying our initial conditions leaves us with
\[
v(y)=c\sin{\sqrt{2}y},
\]
which yields $\sqrt{2}R=\pi$. Then
\[
{1\over\pi}R={1\over\pi}\int_{0}^{\pi}u(x)\dx={\sqrt{2}\over2},
\]
which was the desired result.

\section{Non-Linear Analysis}

We have shown that $u_{s}$ is a unique solution for for the system described in \eqref{ssone}. We now
want to relate $u_{s}$ to \eqref{nlinh}. We make the observation that $u_{s}(x)$ is an upper bound
of the solution to \eqref{nlinh}, called $u(x,t)$ when the initial condition $u_{0}(x)<u_{s}(x)$; by
comparison we have that the solution to \eqref{nlinh} exists globally, and
$0\leq u(x,t)\leq u_{s}(x)$. This solution, however is only stable when the initial condition
satisfies
\begin{equation}\label{bound}
\int_{0}^{\pi}u_{0}(x)\sin{x}\dx\leq2.
\end{equation}
If $u_{0}(x)$ is too large, the solution blows up as $t\to\infty$. We will now try to show that the
solution exits when \eqref{bound} holds.

\subsection{Analysis of the Initial Conditions}
We begin by letting
\[
F(t)=\int_{0}^{\pi}u(x,t)\sin{x}\dx.
\]
Differentiating with respect to $t$ yields
\begin{align*}
F'(t)&=\int_{0}^{\pi}u_{t}\sin{x}\dx\\
&=\int_{0}^{\pi}(u_{xx}+u^{3})\sin{x}\dx\\
&=
\end{align*}
\end{document}
